---
title: "Assignment 2 Submission"
author: "Your name(s)"
date: "submission date"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statement of Contribution

Include this if pairing.

State what each team member has contributed to the assignment before answering the questions. Each student must contribute towards ALL questions. Your assignment will not be marked without this statement.

```{r eval=FALSE, include=FALSE}
## There are a few options in every chunk, click the setting button (greyed) near the triangle:
# 1. to use default as in the above (show code and output, ie ECHO=TRUE) or 
# 2. to show output only (ie ECHO=FALSE) or
# 3. show nothing (run code) ie include=FALSE
# 4. show nothing (dont run the code), like this chunk, ie eval=FALSE, include=FALSE
## You can knit this template into Word and then update it into pdf etc..
## You can create your own way for submission, this is just an example if you are going to use Rmd.
```

# Question 1: Air Pollution and Mortality. Does pollution kill people? (30 marks)

(a) (5 marks) Carry out exploratory data analysis (EDA) of this dataset.

    **Answer:**

    *Answer text here*

    ```{r}
    # Answer code here
    ```

**Continue with other questions here.**

# Question 2: Body Measurements (25 marks)

(a) (4 marks) Carry out exploratory data analysis (EDA) of this dataset before you do any modelling.

    **Answer:**

    *Answer text here*

**Continue with other questions here.**
